---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# datingfourthaisites

[![Circle-CI Build Status](https://circleci.com/gh/benmarwick/14C_Antiquity.svg?style=shield&circle-token=:circle-token)](https://circleci.com/gh/benmarwick/14C_Antiquity)

This repository contains the data and code for our paper:

> Conrad, C., Marwick, B., Higham, C., Castillo, C., Fuller, D., White, J.C., Shoocongdej, R., Thongcharoenchaikit, C., Green, E. and Yen, D.  submitted. Reanalysis of the Radiocarbon and Thermoluminescence Chronologies for Spirit Cave, Steep Cliff Cave, Banyan Valley Cave and Non Nok Tha in Northern Thailand. _Antiquity_ <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Conrad, C., Marwick, B., Higham, C., Castillo, C., Fuller, D., White, J.C., Shoocongdej, R., Thongcharoenchaikit, C., Green, E. and Yen, D., (YYYY). _Reanalysis of the Radiocarbon and Thermoluminescence Chronologies for Spirit Cave, Steep Cliff Cave, Banyan Valley Cave and Non Nok Tha in Northern Thailand_ Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Conrad, C. and B. Marwick (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for 'Reanalysis of the Radiocarbon and Thermoluminescence Chronologies for Spirit Cave, Steep Cliff Cave, Banyan Valley Cave and Non Nok Tha in Northern Thailand'_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: </archive/master.zip>

Or you can install this compendium as an R package, datingfourthaisites, from GitHub with:

```{r eval = FALSE}
source("https://install-github.me/cylerc/datingfourthaisites")
```


### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

