---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# datingfourthaisites

 ```{r echo=FALSE, prompt=FALSE, results='asis', comment=""}
cat('
[![CircleCI](https://circleci.com/gh/benmarwick/datingfourthaisites/tree/master.svg?style=svg)](https://circleci.com/gh/benmarwick/datingfourthaisites/tree/master) [![Binder](http://mybinder.org/badge.svg)](http://beta.mybinder.org/v2/gh/benmarwick/datingfourthaisites/master?urlpath=rstudio)
')
```



This repository contains the data and code for our paper:

> Conrad, C., Shoocondej, R., Marwick, B., White, J.C., Thongcharoenchaikit, C., Higham, C., Feathers, J.K., Castillo, C. and Fuller, D. submitted. Radiocarbon and Luminescence Dating of Spirit Cave, Steep Cliff Cave and Banyan Valley Cave in Northwest Thailand.  _Antiquity_ <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Conrad, C., Shoocondej, R., Marwick, B., White, J.C., Thongcharoenchaikit, C., Higham, C., Feathers, J.K., Castillo, C. and Fuller, D., (YYYY). _Radiocarbon and Luminescence Dating of Spirit Cave, Steep Cliff Cave and Banyan Valley Cave in Northwest Thailand_ Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Conrad, C., B. Marwick and Feathers, J.K. (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for 'Radiocarbon and Luminescence Dating of Spirit Cave, Steep Cliff Cave and Banyan Valley Cave in Northwest Thailand'_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: </archive/master.zip>

Or you can install this compendium as an R package, datingfourthaisites, from GitHub with:

```{r eval = FALSE}
source("https://install-github.me/cylerc/datingfourthaisites")
```


### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

